\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{small} Thanks to Robert Potvliege, who guided me through the whole {PhD},
and taught me so much both about physics and the curiosity and rigour needed to
be a good scientist. It has been a pleasure to work as your student. Thanks to
Charles Adams, who brought both interesting experimental systems for me to
model, and his enthusiam and insight in helping to uncover the underlying
physics. I already miss our conversations about physics and much else, including
on our runs around the Wear.

Thanks to Ifan Hughes, always generous with his time to explain anything from
the nature of refractive index to why Swansea beat Liverpool in the football.
Thanks to Simon Gardiner, whose warm welcome when I first came to visit
encouraged me to come to Durham. I'm very glad I did. Thanks to James Keaveney
and Kate Whittaker, for the excellence in experiment that gave me the data to
work with.

Thanks as a whole to the \textsc{jqc} group, whose winning combination of
excellence and community made the office such a pleasant place to work. Too many
people have contributed to that in my years here to list individually, and I am
sorry to everyone I leave out. But I must give thanks to Danielle Boddy,
my long-suffering running buddy (\ie coach). Thanks for putting up with so much
of my `rambling chat'. Thanks to David Parades, who with his infectious
enthusiasm motivated me often both in physics and to get up and meet him at
the gym. Thanks to Chris Wade, in particular for taking me out on some memorable
bike rides and not leaving me behind, even when I forgot to bring lights. Thanks
to Christophe Valliant, especially for answering my many, probably repeated,
questions about angular momentum theory. Thanks to Hannes Busche for the
great fun we had blasting around Arizona and Utah in the `Canyonero'.

Thanks to those I enjoyed sharing an office and many conversations over tea
with: Tom Billam, Rob Bettles, Michael K\"{o}ppinger, Mark Zentile, John Helm,
Liz Bridge, Anna Marchant; and to the rest of my cohort: Arin Mizuori, Dan
Saddler, Manfred Yu and Alfred Millett-Sikking.

Thanks to the housemates who helped make my years in Durham so enjoyable: Cesca
Richards, Barney Chesterton, Donald MacLennan and Jamie Taylor. It was the best
of times, it was the blurst of times.

Finally thanks to James Keaveney, Tom Billam and Steven Wrathmall for helping me
to dot the $i$'s and cross the $\hbar$'s.
\end{small}
